---
title: "Assignment 4: Exploratory Data Analysis"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Sunday, 9/28 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot & interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project data folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
library(tidyverse)

mydata <- read_csv("data/boardgame-geek-dataset_organized.csv")

```

### Describe your data

Data author: Chikodi Obu (however data was scaped from boardgamegeek.com which originally collected the data

Link: <https://www.kaggle.com/datasets/chik0di/board-games-dataset-complete-features?select=boardgame-geek-dataset_organized.csv>

Description: This dataset contains comprehensive information about board games scraped from BoardGameGeek. It includes game specifications, user ratings, popularity metrics, designer credits, marketplace pricing for games available on amazon.

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}
library(DataExplorer)

# see the table as a whole 
view(mydata) 

# use DataExplorer to see types of columns and missing values
introduce(penguins_data)
plot_intro(penguins_data)

# use summary to see the length and stats of the columns 
summary(mydata) 
```

### Summary stats

Before you make plots, you must also:

-   Show basic summaries of your dataset, including:

    -   `count()` for categorical variables (to see category frequencies)

    -   `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

```{r}

mydata$designers <- as.factor(mydata$designers)

mydata %>% 
  count(designers)

summary(mydata$max_players)
sd(mydata$max_players)

summary(mydata$rating_1)
sd(mydata$rating_1)
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

From attempting to count the categorical variables I discovered that almost every row is unique. From the summaries of the ratings I learned that the ratings are the number of times a game got that rating on a one to ten scale.

## Part 2:

Create **at least 4 plots**, each of a different type. At least one of your four plots must use different variables than the others.

Each plot should have:

-   A title and labeled axes

-   A short interpretation (2-3 sentences) underneath

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can be plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}
ggplot(mydata, aes(x = max_players)) + 
  geom_histogram(binwidth = 1, fill = "#3b7541") 

```

Interpretation: Most games have a max player count that is less than ten, however there are a few games with a max of 100. This is likely due to some games putting the arbitrary limit of 100 players to indicate there is no real limit.

### Plot 2

```{r}
ggplot(mydata, aes(x= complexity, y = avg_rating)) +
  geom_point(color = "#300b40") +
  geom_smooth(method = "lm", se = FALSE, color = "blue")

```

Interpretation: Generally more complex games are better received by their audiences

### Plot 3

```{r}
ggplot(mydata, aes(x = release_year, y = min_playtime))+
  geom_line()


```

Interpretation: Board games have become more varied in playtime in the 21st century. Likely due to mass production and innovation in the industry

### Plot 4

```{r}
ggplot(mydata, aes(x =minimum_age))+
  geom_density()

```

Interpretation: most games require the player to be older than 7 and many require the player to be older than 14

## Push to GitHub to Submit!

Assignment is due Sunday 9/28 at 11:59 pm.
